MSK III

### 2017-07-20 msk3-gift.rmd
Added short case x | Typose in the gift

|  Type                        |  Relative cost (Canada; 2000)  | Radiation           | What can see                                                                           | What can't one see                 |
|------------------------------|--------------------------------|---------------------|----------------------------------------------------------------------------------------|------------------------------------|
|  Plain radiographs           |  1 (hand)                      | Low (in comparison) | Bone                                                                                   | Soft tissues                       |
|  Computed tomography         |  8 (lumbar spine)              | High                | Bone high detail Soft tissues Brain aneurysms Chest, abdo, pelvis                      | Soft tissues not AS good as MRI    |
|  Magnetic resonance imaging  |  18 (knee)                     | None                | Soft tissues amazingly. E.g. ligaments, tendons, spinal cord injury, brain tumours,... | Bone is not as clear as xray or CT |
|  Ultrasound                  |  2 (knee)                      | None                | Soft tissues, joint effusions Intra-abdominal organs Flow of blood Foetus              | Bones                              |
|  Bone scan                   |  7?                            | Medium              | Metabolically active bone                                                              |                                    |

@@

Gold standard of soft tissue imagnig, no radiation required. The most frequent indications for MRI are for diseases of the neurological and musculo-skeletal systems. Other indications include tumour staging (for example rectal, prostatic and gynaecological malignancies), hepatobiliary disease, assessment of the breast in patients of high-risk for cancer, and of breast implants. Increasingly, MRI is substituting for CT scans particularly in children and young adults because of the lack of ionizing radiation. However, in Australia the indications that attract a Medicare rebate are limited.

>>>>

|  Type                        |  Relative cost (Canada; 2000)  | Radiation           | What can see                                                                           | What can't one see                 |
|------------------------------|--------------------------------|---------------------|----------------------------------------------------------------------------------------|------------------------------------|
|  Plain radiographs           |  1 (hand)                      | Low (in comparison) | Bone                                                                                   | Soft tissues                       |
|  Computed tomography         |  8 (lumbar spine)              | High                | Bone high detail. Soft tissues. Brain aneurysms. Chest, abdo, pelvis                      | Soft tissues not AS good as MRI    |
|  Magnetic resonance imaging  |  18 (knee)                     | None                | Soft tissues amazingly. E.g. ligaments, tendons, spinal cord injury, brain tumours,... | Bone is not as clear as xray or CT |
|  Ultrasound                  |  2 (knee)                      | None                | Soft tissues, joint effusions Intra-abdominal organs Flow of blood Foetus              | Bones                              |
|  Bone scan                   |  7?                            | Medium              | Metabolically active bone                                                              |                                    |

@@

Gold standard of soft tissue imaging, no radiation required. The most frequent indications for MRI are for diseases of the neurological and musculo-skeletal systems. Other indications include tumour staging (for example rectal, prostatic and gynaecological malignancies), hepatobiliary disease, assessment of the breast in patients of high-risk for cancer, and of breast implants. Increasingly, MRI is substituting for CT scans particularly in children and young adults because of the lack of ionizing radiation. However, in Australia the indications that attract a Medicare rebate are limited.

### 2017-07-20 msk3.rmd
Added short case x | Typose in the gift

@@

>>>>

Knee structures

-  Bones: patella, tibia, fibula, femur
-  Soft tissue: tendons, menisci, ligaments

Contact sport: ACL (cruciates: snap and then "ow" | non-contact change in direction), lateral collateral ligament, med meniscus

Questions:

-  HPC
-  Did you hit your head? Remember a snap/something? How exactly did you fall? Allergies?

Bucket list for knee pain

-  Acute: trauma
-  Chronic: knee pain without trauma

Ligaments responding to stress

-  Special tests may give *false negatives* straight away

Effusion: in a joint space | potential space

Oedema: outside a joint

Mx of ACL injury: (rupture/tear)

-  Initial: RICE = {rest, ice, compression, elevation}
-  Non-operative: physio, balance
-  Operative: surgical repair

@@

How are hips checked:

-  Barlow: can dislocate the hip
-  Ortolani: pull it out

Galeazzi test: look at height of knee

Tx of DDH:

-  Van Rosen splint: QLD + Scandinavia
-  Pavlik harness
-  If instability at 10 years then surgery

Leg length discrepancy: dislocation (moves posteriorly and shortening)

-  SUFE: 10-16
-  Perthes: 4-10
